### 2024-07-30 getting-started.qmd
20240722 - fix nested list
20240730 - separate weekly and seasonal PFR; compute more player ages from PFR

# Downloaded Data
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_teams.RData")
load(file = "./data/nfl_rosters.RData")
load(file = "./data/nfl_rosters_weekly.RData")
load(file = "./data/nfl_schedules.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/nfl_draftPicks.RData")
load(file = "./data/nfl_depthCharts.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
load(file = "./data/nfl_actualStats_weekly.RData")
load(file = "./data/nfl_injuries.RData")
load(file = "./data/nfl_snapCounts.RData")
load(file = "./data/nfl_espnQBR.RData")
load(file = "./data/nfl_nextGenStats_weekly.RData")
load(file = "./data/nfl_advancedStatsPFR.RData")
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_ftnCharting.RData")
load(file = "./data/nfl_playerIDs.RData")
load(file = "./data/nfl_rankings.RData")
load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
load(file = "./data/nfl_expectedFantasyPoints_offense.RData")

@@

### Advanced Stats from PFR {#sec-advancedStatsPFR}

A Data Dictionary for PFR passing data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pfr_passing.html>

@@

nfl_advancedStatsPFR <- bind_rows(
  nfl_advancedStatsPFR_pass_seasonal,
  nfl_advancedStatsPFR_pass_weekly,
  nfl_advancedStatsPFR_rush_seasonal,
  nfl_advancedStatsPFR_rush_weekly,
  nfl_advancedStatsPFR_rec_seasonal,
  nfl_advancedStatsPFR_rec_weekly,
  nfl_advancedStatsPFR_def_seasonal,
  nfl_advancedStatsPFR_def_weekly,
)
```

@@

save(
  nfl_advancedStatsPFR,
  file = "./data/nfl_advancedStatsPFR.RData"
)
```

@@

load(file = "./data/nfl_advancedStatsPFR.RData")
```

@@

nfl_expectedFantasyPoints_offense <- bind_rows(
  nfl_expectedFantasyPoints_pass,
  nfl_expectedFantasyPoints_rush
)
```

@@

save(
  nfl_expectedFantasyPoints_offense,
  file = "./data/nfl_expectedFantasyPoints_offense.RData"
)
```

@@

load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
load(file = "./data/nfl_expectedFantasyPoints_offense.RData")
```

@@

@@

>>>>

# Downloaded Data
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_teams.RData")
load(file = "./data/nfl_rosters.RData")
load(file = "./data/nfl_rosters_weekly.RData")
load(file = "./data/nfl_schedules.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/nfl_draftPicks.RData")
load(file = "./data/nfl_depthCharts.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
load(file = "./data/nfl_actualStats_weekly.RData")
load(file = "./data/nfl_injuries.RData")
load(file = "./data/nfl_snapCounts.RData")
load(file = "./data/nfl_espnQBR.RData")
load(file = "./data/nfl_nextGenStats_weekly.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_ftnCharting.RData")
load(file = "./data/nfl_playerIDs.RData")
load(file = "./data/nfl_rankings.RData")
load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")

@@

### Advanced Stats from Pro Football Reference {#sec-advancedStatsPFR}

A Data Dictionary for Pro Football Reference passing data is located at the following link: <https://nflreadr.nflverse.com/articles/dictionary_pfr_passing.html>

@@

nfl_advancedStatsPFR_seasonal <- bind_rows(
  nfl_advancedStatsPFR_pass_seasonal,
  nfl_advancedStatsPFR_rush_seasonal,
  nfl_advancedStatsPFR_rec_seasonal,
  nfl_advancedStatsPFR_def_seasonal
)

nfl_advancedStatsPFR_weekly <- bind_rows(
  nfl_advancedStatsPFR_pass_weekly,
  nfl_advancedStatsPFR_rush_weekly,
  nfl_advancedStatsPFR_rec_weekly,
  nfl_advancedStatsPFR_def_weekly
)
```

Merge with `gsis_id` for merging with other datasets:

@@

nfl_advancedStatsPFR_weekly <- nfl_advancedStatsPFR_weekly %>% 
  rename(pfr_id = pfr_player_id)

nfl_advancedStatsPFR_seasonal <- left_join(
  nfl_advancedStatsPFR_seasonal,
  nfl_playerIDs %>% 
    filter(!is.na(pfr_id)) %>% 
    filter(gsis_id != "00-0039137") %>% # drop DL Byron Young, keep OLB Byron Young
    select(pfr_id, gsis_id) %>% 
    unique(),
  by = "pfr_id"
)

nfl_advancedStatsPFR_weekly <- left_join(
  nfl_advancedStatsPFR_weekly,
  nfl_playerIDs %>% 
    filter(!is.na(pfr_id)) %>% 
    filter(gsis_id != "00-0039137") %>% # drop DL Byron Young, keep OLB Byron Young
    select(pfr_id, gsis_id) %>% 
    unique(),
  by = "pfr_id"
)

nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0035665" & nfl_advancedStatsPFR_seasonal$pos %in% c("LB","LILB","RILB"))] <- NA # drop LB David Young, keep DB David Young
#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0035665" & nfl_advancedStatsPFR_weekly$team %in% c("TEN","MIA"))] <- NA # drop LB David Young, keep DB David Young

nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0035292" & nfl_advancedStatsPFR_seasonal$pos %in% c("LB","LILB","RILB"))] <- NA # drop LB David Young, keep DB David Young
nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0035292" & nfl_advancedStatsPFR_weekly$team %in% c("TEN","MIA"))] <- NA # drop LB David Young, keep DB David Young

nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0033894" & nfl_advancedStatsPFR_seasonal$pos == "DB")] <- NA # drop S Marcus Williams, keep DB David Young
#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0033894" & nfl_advancedStatsPFR_weekly$pos == "DB")] <- NA # drop S Marcus Williams

nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0038407" & nfl_advancedStatsPFR_seasonal$pos == "DB")] <- NA # drop DB Jaylon Jones, keep CB Jaylon Jones
#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0038407" & nfl_advancedStatsPFR_weekly$pos == "DB")] <- NA # drop DB Jaylon Jones, keep CB Jaylon Jones

nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0037106" & nfl_advancedStatsPFR_seasonal$pos == "DB")] <- NA # drop DB Jaylon Jones, keep CB Jaylon Jones
#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0037106" & nfl_advancedStatsPFR_weekly$pos == "DB")] <- NA # drop DB Jaylon Jones, keep CB Jaylon Jones

nfl_advancedStatsPFR_seasonal$gsis_id[which(nfl_advancedStatsPFR_seasonal$gsis_id == "00-0038549" & nfl_advancedStatsPFR_seasonal$pos == "WR")] <- NA # drop WR DJ TUrner, keep CB DJ Turner
#nfl_advancedStatsPFR_weekly$gsis_id[which(nfl_advancedStatsPFR_weekly$gsis_id == "00-0038549" & nfl_advancedStatsPFR_weekly$pos == "WR")] <- NA # drop WR DJ TUrner, keep CB DJ Turner
```

save(
  nfl_advancedStatsPFR_seasonal,
  file = "./data/nfl_advancedStatsPFR_seasonal.RData"
)

save(
  nfl_advancedStatsPFR_weekly,
  file = "./data/nfl_advancedStatsPFR_weekly.RData"
)
```

@@

load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
```

@@

nfl_expectedFantasyPoints_pbp <- bind_rows(
  nfl_expectedFantasyPoints_pass,
  nfl_expectedFantasyPoints_rush
)
```

@@

save(
  nfl_expectedFantasyPoints_pbp,
  file = "./data/nfl_expectedFantasyPoints_pbp.RData"
)
```

@@

load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")
```

@@

# Merge with Pro Football Reference Data on Player Age by Season
player_game_birthdate_gameday_offense <- player_game_birthdate_gameday_offense %>% 
  dplyr::left_join(
    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
    by = c("player_id" = "gsis_id", "season")
  )

player_game_birthdate_gameday_defense <- player_game_birthdate_gameday_defense %>% 
  dplyr::left_join(
    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
    by = c("player_id" = "gsis_id", "season")
  )

player_game_birthdate_gameday_kicking <- player_game_birthdate_gameday_kicking %>% 
  dplyr::left_join(
    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
    by = c("player_id" = "gsis_id", "season")
  )

# Set age as first non-missing value from calculation above or from PFR
player_game_birthdate_gameday_offense <- player_game_birthdate_gameday_offense %>% 
  mutate(age = coalesce(age.x, age.y)) %>% 
  select(-age.x, -age.y)

player_game_birthdate_gameday_defense <- player_game_birthdate_gameday_defense %>% 
  mutate(age = coalesce(age.x, age.y)) %>% 
  select(-age.x, -age.y)

player_game_birthdate_gameday_kicking <- player_game_birthdate_gameday_kicking %>% 
  mutate(age = coalesce(age.x, age.y)) %>% 
  select(-age.x, -age.y)

@@

# Merge with Pro Football Reference Data on Player Age by Season
player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
  dplyr::left_join(
    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
    by = c("player_id" = "gsis_id", "season")
  )

player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
  dplyr::left_join(
    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
    by = c("player_id" = "gsis_id", "season")
  )

player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
  dplyr::left_join(
    nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(age)) %>% select(gsis_id, season, age) %>% unique(),
    by = c("player_id" = "gsis_id", "season")
  )

# Set age as first non-missing value from calculation above or from PFR
player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
  mutate(age = coalesce(age.x, age.y)) %>% 
  select(-age.x, -age.y)

player_stats_seasonal_defense <- player_stats_seasonal_defense %>% 
  mutate(age = coalesce(age.x, age.y)) %>% 
  select(-age.x, -age.y)

player_stats_seasonal_kicking <- player_stats_seasonal_kicking %>% 
  mutate(age = coalesce(age.x, age.y)) %>% 
  select(-age.x, -age.y)
